Volatility Metric to Detect Anomalies in Source Code Repositories

### 2020-04-19 paper/article.tex
more

\documentclass[sigconf]{acmart}
\title{Source Code Volatility Metric}
\author{Yegor Bugayenko}{}{}
\email{yegor.bugayenko@huawei.com}
\affiliation{%
  \institution{Huawei Technologies Co., Ltd.}
  \city{Moscow, Russia}
}
\ccsdesc[100]{Software and its engineering~Metrics}
\keywords{Metrics, Software Size, Software Maintainability}
\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}

@@

Most software development projects keep their source code in Git~\citet{loeliger2012},
which is the de-facto standard in the industry at the moment, or similar
version control systems. Every system, including open
source products like Git, Subversion, and Mercurial, and commerical tools
like Borland StarTeam\texttrademark{} or IBM ClearCase\texttrademark{}
have the same feature: keeping track of the changes
made to the source code files, known as ``logs.''

Git logs provide information about every single change made by every software
developer during the entire course of the project. Using this information
it's possible to measure which files are being modified frequently. On the
other hand, it's also possible to spot files that are rarely modified and may
be considered as ``dead'' or ``abandoned'' code, which is a threat
to the maintainability of the entire project.

It is possible to introduce a metric to measure the relationship
between the amount of actively modified files and files which stay
in the repository for a long time without modifications.
In Section~\ref{sec:method} such a metric is introduced.

In order to demonstrate how the metric works we applied it to
\thetotalrepos{} public Java repositories from GitHub and analyzed
its impact on a few other simple metrics, including the number
of files in a repository, the number of directories, and the amount
of bytes. Since, according to~\citet{muthukumaran2015},
``code change metrics mined from source control repositories have
proven to be the most reliable predictors of bugs in
contemporary software engineering research,'' the introduced volatility
metric may be used to spot maintainability issues.

@@

The metric introduced in this research belongs to the family of
``code change metrics,''
Code churn is one of the simplest metric in the family, which counts
all lines being modified (added, edited, and deleted) during
the entire lifetime of a project~\citep{munson1998}. It was demonstrated
by~\citet{shin2010} how code churn, together with complexity and other
developer activity metrics, is related to software vulnerabilities.
Yet another simple metrict from the family is
the number of changes being made to a particular file, class, method
or line of code is yet another simple metric, used for example by~\citet{demeyer2000}
in order to analyze the effect of refactorings.

@@

\section{Method}
\label{sec:method}

@@

Then, the entire interval between $\check{T}$ (the maximum value)
and $\hat{T}$ (the minimum value) is divided to $Z$ equivalent groups:

>>>>

\documentclass[sigconf]{acmart}
\title{Volatility Metric to Detect Anomalies in Source Code Repositories}
\author{Yegor Bugayenko}{}{}
\email{yegor.bugayenko@huawei.com}
\affiliation{%
  \institution{Huawei Technologies Co., Ltd.}
  \city{Moscow, Russia}
}
\ccsdesc[100]{Software and its engineering~Metrics}
\keywords{Metrics, Software Size, Software Maintainability}
\setcopyright{acmcopyright}
\copyrightyear{2020}
\acmYear{2020}

@@

Most software development projects keep their source code in Git~\citep{loeliger2012},
which is the de-facto standard in the industry at the moment, or similar
version control systems. Every system, including open
source products like Git, Subversion, and Mercurial, and commerical tools
like Borland StarTeam\texttrademark{} or IBM ClearCase\texttrademark{}
have the same feature: keeping track of the changes
made to the source code files, also known as ``logs.''

Git logs provide information about every single change made by every software
developer during the entire course of the project. Using this information
it's possible to measure which files are being modified frequently. On the
other hand, it's also possible to spot files that are rarely modified and may
be considered as ``dead'' or ``abandoned'' code, which may be considered as a threat
to the maintainability of the entire project.

We introduce a metric to measure the relationship
between the amount of actively modified files and files which stay
in the repository for a long time without modifications.

In order to demonstrate how the metric works we applied it to
\thetotalrepos{} public Java repositories from GitHub and analyzed
its impact on a two other characteristics of each repository:
the number of files in a repository and the size of it in bytes.
Larger repositories shown larger values of the volatility.

@@

``Code change metrics mined from source control repositories have
proven to be the most reliable predictors of bugs in
contemporary software engineering research,'' says \citet{muthukumaran2015}.

Code churn is one of the simplest metric in the family, which counts
all lines being modified (added, edited, and deleted) during
the entire lifetime of a project~\citep{munson1998}. It was demonstrated
by~\citet{shin2010} how code churn, together with complexity and other
developer activity metrics, is related to software vulnerabilities.
Yet another simple metrict from the family is
the number of changes being made to a particular file, class, method
or line of code is yet another simple metric, used for example by~\citet{demeyer2000}
in order to analyze the effect of refactorings.

@@

\section{Metric}
\label{sec:method}

@@

Then, the entire interval between $\check{T}$ (the maximum value)
and $\hat{T}$ (the minimum value) is divided to $Z$ (the input parameter of the method)
equivalent groups:
